---
title: "Opal workflow"
author: "Lars van der Burg"
date: "2024-02-27"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(opalr)
library(tidyverse)
library(diffdf)
library(openxlsx)
library(rtres)

library(OpalLUMCADM)
```



Here a Rmarkdown to show all the steps in a general workflow for working with Opal. So from creating a table, moving it, adjusting it and deleting it. Everything will be explained via one of two datasets (you can specify dataset = 1 or 2 below)

# Initializations
```{r}
dataset = 1  # With which dataset do you want to run everything? 1 or 2?
projname = "TEST_LARS"

path_diffdf = paste0(getwd(), "/Diffdf/")
if(!dir.exists(path_diffdf)){
  dir.create(path_diffdf)
}

opal_test <- opalr::opal.login(url = "https://dw-test.clinicalresearch.nl/repo",
                               token = keyring::key_get("token_opal_testclinicalresearch"))
```


# Dataset
```{r}
if(dataset == 1){
  # Random data test file:
  load("FAKE_datafile.RData") |> as_tibble()
  load("FAKE_var.RData") |> as_tibble()
  load("FAKE_cat.RData") |> as_tibble()
  

  tablename = "TESTTABLE1"
  ent = "R"
  
  
  datafiles = list("first" = datafile, "second" = datafile, "third" = datafile) 
  vars = list("first" = var, "second" = var, "third" = var) 
  cats = list("first" = cat, "second" = cat, "third" = cat)
  
  tablenames = c("TESTTABLE1", "TESTTABLE1", "TESTTABLE1")
  tablenames2 = c("TESTTABLE1_copy1", "TESTTABLE1_copy2", "TESTTABLE1_copy3")
  
  
} else if(dataset == 2){

  # Random data test file:
  load(file = "randomdata_removed-parts-importscript_datafile.Rdata")
  datafile = datafile |> as_tibble()
  
  # Load dictionary file(s) into R. At least include a variables dictionary. categories dict is optional.
  load(file = "randomdata_removed-parts-importscript_var.Rdata")
  load(file = "randomdata_removed-parts-importscript_cat.Rdata")
  
  var = var |> as_tibble()
  cat = cat |> as_tibble()
  
  
  ## Can't use datafile_conform_var because there are purposely made mistakes in the datafile
  # datafile = datafile_conform_var(datafile = datafile, var = var)$datafile
  # datafile = opalr::dictionary.apply(tibble = datafile, variables = var, cat = cat)
  

  tablename = "TESTTABLE2"
  ent = "L"

  
  datafiles = list("first" = datafile, "second" = datafile, "third" = datafile) 
  vars = list("first" = var, "second" = var, "third" = var) 
  cats = list("first" = cat, "second" = cat, "third" = cat)
  
  tablenames = c("TESTTABLE2", "TESTTABLE2", "TESTTABLE2")
  tablenames2 = c("TESTTABLE2_copy1", "TESTTABLE2_copy2", "TESTTABLE2_copy3")

  
} else {
  stop("Choose dataset 1 or 2, or supply elsewhere")
}
```



# Write table
Basis function just to write a table from R to opal.
```{r}
var = var |> mutate(encrypted = rep(c("yes", "no", "SI", "YES", "Si", "No", "NO", "SI", "yes", "no"), 2))

write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis"), 
                   datafile = datafile, var = var, cat = cat, ent = ent, 
                   action = "write", id.name = "id", max_tries = 2)
```

## Timezone 
```{r}
datafile$datetime1a = datafile$datetime1
datafile$datetime1b = datafile$datetime1

datafile$datetime1a = with_tz(datafile$datetime1a, "Europe/Amsterdam"); attributes(datafile$datetime1a)$tzone = "Europe/Amsterdam"
datafile$datetime1b = with_tz(datafile$datetime1b, "UTC"); attributes(datafile$datetime1b)$tzone = "UTC"


write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis_tz"), 
                   datafile = datafile, var = var, cat = NULL, ent = ent, 
                   action = "write", id.name = "id", max_tries = 2)
``` 


# Read table
Basis function just to read in a table from opal and make it R ready by discarding some of the attributes
```{r}
imported_table = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis"), id.name = "id", max_tries = 2)
imported_table_tz = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis_tz"), id.name = "id", max_tries = 2)
```


# diffdf report
Function to obtain the differences between the datafiles and dictionaries. There are many different ways this function can be used, check the function documentation.
```{r}
datafile2 = imported_table$datafile
var2 = imported_table$var
cat2 = imported_table$cat


datafile2[1, "text1"] = " 4RRnDdnOBEYa8qDTrNSjvMyu0"
datafile2[2, "text1"] = "KfQs5tyxEsgGeuMqfCFgXZR66 "
datafile2[3, "text1"] = " T8uIXL8xc8ZjeX2VvVksUvr "
  
report_diffdf = check_diffdf_opal_generic(datafile = datafile, datafile2 = datafile2, var = var, var2 = var2, cat = cat, cat2 = cat2, 
                                          comparison = "mod", comp_key = "id", date_time_format = c("Ymd", "Ymd HMS"), suppress_warnings = TRUE, 
                                          report_path = path_diffdf, report_name = "Report_diffdf", report_title = "This is a trial run",
                                          opt_repl_castor = TRUE, opt_rm_VarDiff_null = FALSE, aggregate_VarDiff = TRUE, 
                                          opt_calc_VarDiff_spaces  = TRUE, opt_calc_VarDiff_diff = TRUE, opt_rm_VarDiff_diff_0 = FALSE) 


report_diffdf$VarDiff
report_diffdf$VarDiff |> filter(difference == 0)
# report_diffdf$VarDiff_aggregated
```


# Create table
Function to import a table into opal. 

An optional diffdf report will be made if report_path != FALSE. if report_path = NULL, it will not be saved, just outputted by the function
```{r create}
report_import1 <- import_create_table_opal(opal = opal_test, projname = projname, tablename = tablename, ent = ent, 
                                           datafile = datafile, var = var, cat = cat, action = "write", id.name = "id", comparison = "mod",
                                           report_path = path_diffdf, report_name = paste("Report_import", tablename, sep = "_"))


report_import2 <- import_create_table_opal(opal = opal_test, projname = projname, tablename = tablename, ent = ent, 
                                           datafile = datafile, var = var, cat = cat, action = "update", id.name = "id", comparison = "mod",
                                           report_path = path_diffdf, report_name = paste("Report_import_update", tablename, sep = "_"))


report_import3 <- import_create_table_opal(opal = opal_test, projname = projname, tablename = tablename, ent = ent, 
                                           datafile = datafile, var = var, cat = cat, action = "overwrite", id.name = "id", comparison = "mod",
                                           report_path = path_diffdf, report_name = paste("Report_import_overwrite", tablename, sep = "_"))
```



# Move/copy table
Function to move a table in opal.

An optional diffdf report will be made if report_path != FALSE. if report_path = NULL, it will not be saved, just outputted by the function
```{r copy}
report_copy <- import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, ent = ent,
                                      opal2 = opal_test, projname2 = projname, tablename2 = paste0(tablename, "_copy"), 
                                      report_path = path_diffdf, report_name = paste("Report_copy", tablename, sep = "_"))
```

Possible to move more than table at once, with a wrapper around `import_copy_table_opal`.
```{r multi copy}
report_copy = import_copy_table_opal_many(opal = opal_test, projnames = projname, tablenames = tablenames, ent = ent,
                                          opal2 = opal_test, projnames2 = projname, tablenames2 = tablenames2,
                                          report_path = path_diffdf, report_name = "Report_copy")
```


# Checks
## All checks
There is a single function to run all checks that we currently want to run. It does not change the data, but just gives you an indication whether or not something MIGHT be wrong. Sometimes things are missing for a reason and is a reported issue not a real problem. Always be alert.
```{r}
report_checks = checks_opal_R(datafile = datafile, var = var, cat = cat, key = "id", min_max = TRUE, silent = FALSE)
```


### min/max check
One of the checks that is done in the `checks_opal_R` function is check_categoriesminmax_generic. It is also possible to run this check separately and to run it for multiple datasets at once (that is not possible with the `checks_opal_R` function, that can only handle a single datafile/dictionary)
```{r}
report_minmax = check_categoriesminmax_generic(datafiles = list("test" = datafile), vars = list("test" = var), cats = list("test" = cat), report_path = NULL)

report_minmax_mult = check_categoriesminmax_generic(datafiles = datafiles, vars = vars, cats = cats, report_path = NULL)
```

#### Conditional
With the optional conditional argument one can indicate that min/max should only be checked if a specific condition is met. In the simple example below, there are in variable `Cond2` two variables that are outside the min-max range (id 2 and 4). However, only id 2 meets the required `Cond1 = 1`. If one wants to have this, you can use the `conditional` argument.
```{r}
datafile_cond = bind_cols(id = 1:5, Cond1 = c(1, 1, 1, 0, 0), Cond2 = c(4, 12, NA, 14, NA))
var_cond = bind_cols(name = c("Cond1", "Cond2"), valueType = "integer", `label:en` = c("Cond1 label", "Cond2 label"), index = c(6, 7), min = "0", max = "10")


report_minmax = check_categoriesminmax_generic(datafiles = list("test" = datafile_cond), vars = list("test" = var_cond), report_path = NULL, 
                                               conditional = NULL, includeNA = TRUE)

report_minmax_CondF = check_categoriesminmax_generic(datafiles = list("test" = datafile_cond), vars = list("test" = var_cond), report_path = NULL, 
                                                     conditional = list("Cond2 ~ Cond1 == 1"), includeNA = FALSE)
report_minmax_CondT = check_categoriesminmax_generic(datafiles = list("test" = datafile_cond), vars = list("test" = var_cond), report_path = NULL, 
                                                     conditional = list("Cond2 ~ Cond1 == 1"), includeNA = TRUE)
```


### Datafile conform var
The `checks_opal_R` function also checks if the specified valueTypes are in concordance with the datafile types. Two functions are of interest here: 
1) datafile_conform_var_check, which just signals whether or not a mismatch is detected (this is run inside `checks_opal_R`) 
2) datafile_conform_var_change, which actually changes the data classes of the datafile for such a mismatch (this is NOT run inside `checks_opal_R`). 

So if you want to actually change data classes, you have to run `datafile_conform_var_change` separately. 

#### Check
This function will check two things and give a report with the findings
1) whether the valueTypes are named as Opal wants them
2) whether the class of datafile are compatible with the specified valueType. Conversions can be performed with `datafile_conform_var_change`
```{r}
valueType_check = datafile_conform_var_check(datafile, var)
```


#### Change
This function will convert the datafile class to match with the valueType. If datafile class is a character and valueType is a date or datetime, there is no permanant conversion performed, but we will check the conversion that will be done in opal via the function `lubridate::parse_date_time`. With date_time_format you can specify the formats that are observed in the datafile. 

For dataset == 2: when "Ymd HM" is not supplied, there will be loss of data.
```{r}
datafile_conversion = datafile_conform_var_change(datafile = datafile, var = var, date_time_format = c("Ymd", "Ymd HMS", "Ymd HM"))
datafile = datafile_conversion$datafile


opaltable = opalr::dictionary.apply(tibble = datafile, variables = var, cat = cat)
```


# Views
The idea behind opal is that instead of sharing your actual datafile, you can share a view with others. With `make_opal_view` you can make a view.
```{r}
view_report = make_opal_view(opal = opal_test, projname = projname, tablename = tablename, 
                             opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_view"),
                             var = var, cat = cat, update = FALSE, comp_key = "id", ent = ent, 
                             report_path = FALSE)
```


## View update
The idea is that you can simply change the directory, and than update the view. That is also possible with `make_opal_view`.

Currently, instead of updating, we replace the view. Because with an update it is not yet possible to downsize the view.
```{r}
make_opal_view(opal = opal_test, projname = projname, tablename = tablename, 
               opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_view2"),
               var = var, cat = cat, update = FALSE, comp_key = "id", ent = ent)


var_adj = var; var_adj[var_adj$name == "integer2", "label:en"] = "integer with categories; but with a more fancy label"
cat_adj = cat; cat_adj[cat_adj$variable == "integer2" & cat_adj$`label:en` == "level 1", "label:en"] = "level 1; but fancier"

view_report_update = make_opal_view(opal = opal_test, projname = projname, tablename = tablename, 
                                    opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_view2"),
                                    var = var_adj, cat = cat_adj, update = TRUE, comp_key = "id", ent = ent, report_path = NULL)
```

### Deprecate view
When we discard a variable that is in the view from the table, the view is not accessable anymore
```{r}
report_import4 <- import_create_table_opal(opal = opal_test, projname = projname, tablename = tablename, ent = ent, 
                                           datafile = datafile|> select(-integer1), 
                                           var = var |> filter(name != "integer1"), 
                                           cat = cat, action = "overwrite", id.name = "id", comparison = "mod",
                                           report_path = path_diffdf, report_name = paste("Report_import_overwrite", tablename, sep = "_"))

## Is there an easy way to see if this view is still usable?
```


### Less variables
```{r, eval = FALSE, echo = FALSE}
write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis2"), 
                   datafile = datafile, var = var, cat = cat, ent = ent, force = FALSE)

make_opal_view(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis2"), 
               opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_basis2_view"),
               var = var, cat = cat, update = FALSE, comp_key = "id", ent = ent)
opalr::opal.table_perm_add(opal_test, projname, paste0(tablename, "_basis2_view"), subject = "rmwissels", type = "user", permission = "view")
opalr::opal.table_perm(opal_test, projname, paste0(tablename, "_basis2_view"))


var_adj = var; var_adj = var_adj[var_adj$name != "var5", ]
cat_adj = cat; cat_adj = cat_adj[cat_adj$variable != "var5", ]


make_opal_view(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis2"), 
               opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_basis2_view2"),
               var = var_adj, cat = cat_adj, update = FALSE, comp_key = "id", ent = ent)
opalr::opal.table_perm(opal_test, projname, paste0(tablename, "_basis2_view2"))


make_opal_view(opal = opal_test, projname = projname, tablename = paste0(tablename, "_basis2"), 
               opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_basis2_view"),
               var = var_adj, cat = cat_adj, update = TRUE, comp_key = "id", ent = ent)
opalr::opal.table_perm(opal_test, projname, paste0(tablename, "_basis2_view"))
```


## Different $('script')
It is also possible to give more difficult entity Filters to your view, to alter your data or to just give a subset of your data. 

They are not really investigated and (at the moment) cannot be added to `make_opal_view`.

### Setting up data
```{r}
datafile_adj = datafile; var_adj = var; cat_adj = cat


## 4 decimals
datafile_adj[1, 3] = 8.36545894584682
datafile_adj$var2 = as.character(datafile_adj$var2); var_adj[2, "valueType"] = "text"

## No NULL
datafile_adj[3, 3] = "34"; datafile_adj[4, 3] = "99"

## Date same
# datafile_adj[4, 5] = as.character(as.Date(parse_date_time("20230712", "Ymd")))


write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_EntFilt"),
                   datafile = datafile_adj, var = var_adj, cat = cat_adj, ent = ent)


make_opal_view(opal = opal_test, projname = projname, tablename = paste0(tablename, "_EntFilt"), 
               opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_EntFilt_view"),
               var = var, cat = cat, update = FALSE, comp_key = "id", ent = ent)
```

### Adjustments
```{r}
if(dataset == 1){
  
} else if(dataset == 2){
  
  ## More difficicult sources
  ## https://opaldoc.obiba.org/en/latest/magma-user-guide/methods.html
  new_rows = tribble(
    ~name, ~valueType, ~`label:en`, ~unit, ~index, ~min, ~max, ~script,
    "var6", "integer", "duplicate of var1", "", 6, "", "", "$('var1')",
    "var7", "integer", "var1 + 10", "", 7, "", "", "$('var1').plus(10)",
    "var8", "decimal", "var1 * var1", "", 8, "", "", "$('var1').multiply($('var1'))",
    # "var9", "integer", "var7 / var1", "", 9, "", "", "$('var7').div($('var1'))",  ## ERROR IN OPAL: No such variable var7 --> Dus waarschijnlijk kan die alleen berekeningen doen op datafile variabele die bestaan? Of er is een volgorde? Can't get the view to R anymore
    "var10", "text", "var3 not", "", 10, "", "", "$('var3').not('test')",
    "var11", "text", "var1 vs var2", "", 11, "", "", "if($('var1').lt($('var2')).value()){true}else{false}",
    "var12", "integer", "var1 vs var2", "", 13, "", "", "if($('var1').lt($('var2')).value()){1}else{0}",
    "var13", "text", "var1 vs var2", "", 14, "", "", "if($('var1').lt($('var2')).value()){$('var2')}else if($('var1').gt($('var2')).value()){$('var1')}",  ## Var1 < var2 --> var2; var1 > var2 --> var1
    "var14", "text", "isNumeric", "", 15, "", "", "$var('var1').isNumeric()",  # without var --> ERROR IN OPAL: Cannot find function isNumeric
    "var15", "decimal", "type", "", 16, "", "", "$('var2').type('decimal')",
    "var16", "decimal", "type + 10", "", 17, "", "", "$('var2').type('decimal').plus(10)",
    "var17", "text", "Date diff", "", 18, "", "", "$('var4').eq($('var5'))"
  )
}

var_adj2 = var |>
  mutate(script = paste0("$('", name, "')")) |> 
  bind_rows(new_rows)

cat_adj2 = cat


view_report_update = make_opal_view(opal = opal_test, projname = projname, tablename = paste0(tablename, "_EntFilt"), 
                                    opal_view = opal_test, projname_view = projname, tablename_view = paste0(tablename, "_EntFilt_view2"),
                                    var = var_adj2, cat = cat_adj2, update = TRUE, comp_key = "id", ent = ent, report_path = NULL)


imported_datafile2 <- import_table_opal2R(opal_test, projname, paste0(tablename, "_EntFilt_view2"))
imported_datafile2$datafile
```



# Delete table
Function to easily delete a table from opal
```{r}
delete_table_opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_copy"))
```



# Adjust script
It can happen that you just want to adjust your table or dictionary, and do not want to completely re-do everything. Here is a script exploring some options.

It is not a function. Because the possible adjustments are infinite, it is very difficult to build a comprehensive function for that. But you can simply use the `import_table_opal2R` and `write_table_R2opal` functions. Example is only run with dataset = 2.


Hier bevindingen:
- Rijen kunnen gemakkelijk los toegevoegd worden, een var dictionary is wel essentieel
- Overwrite vs Force. Als dataset al bestaat en wil updaten met nieuwe rij:
  - Overwrite = FALSE & Force = FALSE: error and nothing happens (`Destination table will be updated. There could be data dictionary conflicts. Use 'force' parameter to proceed.`)
  - Overwrite = TRUE & Force = FALSE: error and nothing happens (`Destination table needs to be deleted or truncated. Use 'force' parameter to proceed.`)
  - Overwrite = FALSE & Force = TRUE: updated table: example dataset including the extra row
  - Overwrite = TRUE & Force = TRUE: first delete table and make new table: example dataset only containing the extra row)
  - Conclusie is dus als volgt: overwrite bepaalt wat er gebeurt (FALSE = update; TRUE = overwrite), maar de overwrite actie wordt alleen uitgevoerd als de force parameter = TRUE. Wanneer force = FALSE kan je niet updaten of overwriten.
- Kolommen kunnen ook toegevoegd worden, de var dictionary mag dan wel alleen informatie over de nieuwe kolom bevatten
- Waardes kunnen dan weer wel los aanpepast worden, alleen id en kolom met waarde nodig
- Dictionary kan upgedate worden, entityType in var dictionary is essentieel (anders error (als entityType niet de default `Participant` is))

## Data from opal
```{r}
## To make a separate table for each setting
report_start <- import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename,
                                       projname2 = projname, tablename2 = paste0(tablename, "_adjust"))

import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_adjust_row"))
import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_adjust_col"))
import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_adjust_val"))
import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_remove_val"))
import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_adjust_cat"))



imported_table = import_table_opal2R(opal = opal_test, projname, paste0(tablename, "_adjust"))

datafile = imported_table$datafile
var = imported_table$var
cat = imported_table$cat

datafile4copy = imported_table$datafile4copy
var4copy = imported_table$var4copy
cat4copy = imported_table$cat4copy


if(dataset == 1){

## Add a new row
  
## Add a new column
  
## Due to checking labjournal a missing value can be replaced
  datafile4copy = datafile4copy |> mutate(InjuryHx.InjuryHxID = if_else(id == "Sample1", 418, InjuryHx.InjuryHxID))
  datafile4copy = datafile4copy |> mutate(InjuryHx.GCSPreHospBestEyes = if_else(id == "Sample10", "UN", InjuryHx.GCSPreHospBestEyes))

## New row to cat dictionary  
  
  
} else if(dataset == 2){
  
## Add a new row 
  new_row = tibble::tribble(
    ~id, ~var1, ~var2, ~var3,          ~var4,        ~var5,
    "6", -12L,  -98,   "hypothyroïde", "2002-09-13", "2018-10-01 23:59"
  )
  datafile4copy = datafile4copy |> bind_rows(new_row)
  
  write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_row"), ent = ent,
                     datafile = new_row, var = imported_table$var4copy, cat = imported_table$cat4copy, force = TRUE)
  imported_table2_R = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_row"))


    
  ## Add a new column
  new_col = matrix(c(18.0, 22.5, 23.0, 21.3, 21.2, 28.0) , ncol = 1, dimnames = list(NULL, "var6")) |> as_tibble()
  datafile4copy = datafile4copy |> bind_cols(new_col)
  
  new_col_var = cbind.data.frame("var6", "decimal", NA, FALSE, 6, "BMI", "10", "100") |> as_tibble() |> rename_with(~ colnames(var4copy))
  var4copy = var4copy |> bind_rows(new_col_var)
  
  write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_col"), ent = ent,
                     datafile = datafile4copy |> filter(id < "6") |> select(id, var6), 
                     var = var4copy |> filter(name == "var6"), cat = imported_table$cat4copy, force = TRUE)
  imported_table2_C = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_col"))
  
  
  
  ## Due to checking labjournal a missing value can be replaced
  datafile4copy = datafile4copy |> mutate(var2 = if_else(id == 4, 79, var2))
  datafile4copy = datafile4copy |> mutate(var3 = if_else(id == 2, NA, var3))
  
  write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_val"), ent = ent,
                     datafile = datafile4copy |> mutate(var2 = if_else(id == 4, 79, var2)) |> filter(id == "4") |> select(id, var2),
                     var = var4copy |> filter(name != "var6"), cat = imported_table$cat4copy, force = TRUE)
  imported_table2_V1 = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust_val"))
  
  write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_remove_val"), ent = ent,
                     datafile = datafile4copy |> mutate(var3 = if_else(id == 3, NA, var3)) |> filter(id == "3") |> select(id, var3),
                     var = var4copy |> filter(name != "var6"), cat = imported_table$cat4copy, force = TRUE)
  imported_table2_V2 = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_remove_val"))
  
  
  
  ## New row to cat dictionary  
  new_cat_row = cbind.data.frame(variable = "var1", name = "-98", missing = TRUE, `label:en` = "Not applicable") |> as_tibble()
  cat4copy = cat4copy |> bind_rows(new_cat_row)
  
  opal.table_dictionary_update(opal = opal_test, project = projname, table = paste0(tablename, "_adjust_cat"), 
                               variables = var4copy |> filter(name != "var6") |> mutate(entityType = "L"), categories = cat4copy)
} 


write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust"), ent = ent,
                   datafile = datafile4copy, var = var4copy, cat = cat4copy, force = TRUE)

imported_table2 = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust"))

datafile2 = imported_table2$datafile
var2 = imported_table2$var
cat2 = imported_table2$cat


report_adjust_opal = check_diffdf_opal_generic(datafile = datafile, datafile2 = datafile2, var = var, var2 = var2, cat = cat, cat2 = cat2, 
                                               comparison = "mod", 
                                               report_path = NULL, suppress_warnings = TRUE)


# delete_table_opal(opal = opal_test, projname, paste0(tablename, "_adjust"))
```


## Data from file
```{r}
## To make a separete table on opal that I can adjust
report_start <- import_copy_table_opal(opal = opal_test, projname = projname, tablename = tablename,
                                       projname2 = projname, tablename2 = paste0(tablename, "_adjust2"))


imported_table = import_table_opal2R(opal = opal_test, projname, paste0(tablename, "_adjust2"))

datafile = imported_table$datafile
var = imported_table$var
cat = imported_table$cat


if(dataset == 1){

  
  
} else if(dataset == 2){
## TODO: what are the differences between read in via csv and import_table$datafile4copy. Would say that those should be the same.

path = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\"

datafile4copy = read_csv(paste0(path, "randomdata_removed-parts-importscript_datafile_adjust.csv"), col_names = TRUE, guess_max = 99999, show_col_types = FALSE)
var4copy = read_csv(paste0(path, "randomdata_removed-parts-importscript_var_adjust.csv"), col_names = TRUE, guess_max = 99999, show_col_types = FALSE)
cat4copy = read_csv(paste0(path, "randomdata_removed-parts-importscript_cat_adjust.csv"), col_names = TRUE, guess_max = 99999, show_col_types = FALSE)


datafile_conform_var_check(datafile4copy, var4copy)

datafile4copy_conversion = datafile_conform_var_change(datafile = datafile4copy, var = var4copy, date_time_format = c("Ymd HMS", "Ymd"))
datafile4copy = datafile4copy_conversion$datafile
}


write_table_R2opal(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust2"), ent = ent,
                   datafile = datafile4copy, var = var4copy, cat = cat4copy, force = TRUE)
imported_table2 = import_table_opal2R(opal = opal_test, projname = projname, tablename = paste0(tablename, "_adjust2"))

datafile2 = imported_table2$datafile
var2 = imported_table2$var
cat2 = imported_table2$cat


report_adjust_file = check_diffdf_opal_generic(datafile = datafile, datafile2 = datafile2, var = var, var2 = var2, cat = cat, cat2 = cat2, 
                                               report_path = NULL, suppress_warnings = TRUE)


## To delete the separately made table on opal
# delete_table_opal(opal = opal_test, projname, paste0(tablename, "_adjust2"))
```


# Encryption & Decryption
```{r}
source("C:/Working projects/ADM/C_OpalLUMCADM/R/func_encrypt_data.R")
source("C:/Working projects/ADM/C_OpalLUMCADM/R/func_decrypt_data.R")



vars_to_encrypt = vars_to_decrypt = c("integer1", "text2", "date4")

con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))


TEST = encrypt_data(con = con, data = datafile$id, vars_to_encrypt = "id", search_image = FALSE, salted_encryption = FALSE)
TEST = encrypt_data(con = con, data = datafile$integer1, vars_to_encrypt = "id", search_image = FALSE, salted_encryption = FALSE)


TEST = encrypt_data(con = con, data = datafile, vars_to_encrypt = "id", search_image = FALSE, salted_encryption = FALSE)
TEST = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = FALSE, salted_encryption = FALSE)
TEST = encrypt_data(con = con, data = datafile, vars_to_encrypt = "id", search_image = FALSE, salted_encryption = FALSE)


## No search_image
datafile_encr = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = FALSE, salted_encryption = FALSE)
datafile_decr = decrypt_data(con = con, data = datafile_encr, vars_to_decrypt = vars_to_decrypt)
identical(datafile, datafile_decr |> mutate(integer1 = as.integer(integer1)))


datafile_encr = datafile |> mutate(across(all_of(vars_to_encrypt), ~encrypt_data(con, .x)))
datafile_decr = datafile_encr |> mutate(across(all_of(vars_to_decrypt), ~decrypt_data(con, .x)))
identical(datafile, datafile_decr |> mutate(integer1 = as.integer(integer1)))


datafile_encr = datafile |> 
  mutate(integer1 = encrypt_data(con, datafile$integer1),
         text2 = encrypt_data(con, datafile$text2),
         date4 = encrypt_data(con, datafile$date4))
datafile_decr = datafile_encr |>
  mutate(integer1 = as.integer(decrypt_data(con, datafile_encr$integer1)),
         text2 = decrypt_data(con, datafile_encr$text2),
         date4 = decrypt_data(con, datafile_encr$date4))
all.equal(datafile, datafile_decr)


## Search_image
datafile_encr = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)
datafile_decr = decrypt_data(con = con, data = datafile_encr, vars_to_decrypt = vars_to_decrypt)
```


## Encryption
```{r}
vars_to_encrypt = c("integer1", "text2", "date4")


## Correct connection
con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))

datafile_encr_vec = encrypt_data(con = con, data = datafile$integer1, search_image = TRUE, salted_encryption = FALSE)
datafile_encr_mat = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)


## vars_to_encrypt
datafile_encr_mat1 = encrypt_data(con = con, data = datafile, vars_to_encrypt = NULL, search_image = TRUE, salted_encryption = FALSE)
encrypt_data(con = con, data = datafile, vars_to_encrypt = c("integer0"), search_image = TRUE, salted_encryption = FALSE)



## search_image vs salted_encryption
datafile_encr1 = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)
datafile_encr1 |> mutate(across(all_of(vars_to_encrypt), ~vec_extract_search_image(.x), .names = "{.col}_SI")) |> select(vars_to_encrypt, "integer1_SI", "text2_SI", "date4_SI")

datafile_encr2 = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = FALSE, salted_encryption = TRUE)
datafile_encr2 |> mutate(across(all_of(vars_to_encrypt), ~vec_extract_search_image(.x), .names = "{.col}_SI")) |> select(vars_to_encrypt, "integer1_SI", "text2_SI", "date4_SI")

datafile_encr3 = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = FALSE, salted_encryption = FALSE)
datafile_encr3 |> mutate(across(all_of(vars_to_encrypt), ~vec_extract_search_image(.x), .names = "{.col}_SI")) |> select(vars_to_encrypt, "integer1_SI", "text2_SI", "date4_SI")

datafile_encr4 = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = TRUE)



## Wrong connection
con <- tres_connect(base_url = "something", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))
encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)

con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "something", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))
encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)

con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "something", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))
encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)

con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "something", password = keyring::key_get("TRES_opals2s_lumina"))
encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)

con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = "something")
encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)
```

## Decryption
```{r}
vars_to_decrypt = c("integer1", "text2", "date4")


## Correct connection
con <- tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))

datafile_decr_vec = decrypt_data(con, data = datafile_encr_vec)
identical(datafile$integer1, as.integer(datafile_decr_vec))

datafile_decr_mat = decrypt_data(con, data = datafile_encr_mat, vars_to_decrypt = vars_to_decrypt)
identical(datafile, datafile_decr_mat |> mutate(integer1 = as.integer(integer1)))


## vars_to_decrypt
datafile_decr_mat1 = decrypt_data(con, data = datafile_encr_mat, vars_to_decrypt = NULL)
all.equal(datafile, datafile_decr_mat1 |> mutate(integer1 = as.integer(integer1)))

decrypt_data(con, data = datafile_encr_mat, vars_to_decrypt = "integer0")
datafile_decr_mat2 = decrypt_data(con, data = datafile_encr_mat, vars_to_decrypt = "integer1")
all.equal(datafile, datafile_decr_mat2 |> mutate(integer1 = as.integer(integer1)))

decrypt_data(con, data = datafile_encr_mat, vars_to_decrypt = "integer2")



## Partly encrypted
datafile_sub = datafile; datafile_sub[1, "text1"] = "3::NOT"

datafile_sub_vec = decrypt_data(con, data = datafile_sub$text1)
datafile_sub_mat = decrypt_data(con, data = datafile_sub, vars_to_decrypt = NULL)
datafile_sub_mat = decrypt_data(con, data = datafile_sub, vars_to_decrypt = "text1")


## Wrong connection
# Has same checks as for encrypt_data
```


## Different accounts
```{r}
datafile1 = datafile
datafile2 = datafile
datafile3 = datafile

vars_to_encrypt = vars_to_decrypt = c("integer1", "text2", "date4")


## Encryption
con1 <- tres_connect(base_url = "https://mereden.msbi.nl/TRESTEST/", domain = "test.clinicalresearch.nl", project = "S_O_A_C_B_SUBS_", username = "User1Groep1", password = "groep1USER1!")
datafile1_encr = encrypt_data(con1, datafile1, vars_to_encrypt)


con2 <- tres_connect(base_url = "https://mereden.msbi.nl/TRESTEST/", domain = "test.clinicalresearch.nl", project = "S_O_A_C_B_SUBS_", username = "User2Groep1", password = "groep1USER2!")
datafile2_encr = encrypt_data(con2, datafile2, vars_to_encrypt, search_image = FALSE)


con3 <- tres_connect(base_url = "https://mereden.msbi.nl/TRESTEST/", domain = "test.clinicalresearch.nl", project = "S_O_A_C_B_SUBS_", username = "User3Groep2", password = "groep2USER3!")
datafile3_encr = encrypt_data(con3, datafile3, vars_to_encrypt)



## Decryption
datafile1_decr1 = decrypt_data(con1, datafile1_encr, vars_to_decrypt); identical(datafile1, datafile1_decr1 |> mutate(integer1 = as.integer(integer1)))
datafile2_decr1 = decrypt_data(con1, datafile2_encr, vars_to_decrypt); identical(datafile2, datafile2_decr1 |> mutate(integer1 = as.integer(integer1)))
datafile3_decr1 = decrypt_data(con1, datafile3_encr, vars_to_decrypt); identical(datafile3, datafile3_decr1 |> mutate(integer1 = as.integer(integer1)))

datafile1_decr2 = decrypt_data(con2, datafile1_encr, vars_to_decrypt); identical(datafile1, datafile1_decr2 |> mutate(integer1 = as.integer(integer1)))
datafile2_decr2 = decrypt_data(con2, datafile2_encr, vars_to_decrypt); identical(datafile2, datafile2_decr2 |> mutate(integer1 = as.integer(integer1)))
datafile3_decr2 = decrypt_data(con2, datafile3_encr, vars_to_decrypt); identical(datafile3, datafile3_decr2 |> mutate(integer1 = as.integer(integer1)))

datafile1_decr3 = decrypt_data(con3, datafile1_encr, vars_to_decrypt); identical(datafile1, datafile1_decr3 |> mutate(integer1 = as.integer(integer1)))
datafile2_decr3 = decrypt_data(con3, datafile2_encr, vars_to_decrypt); identical(datafile2, datafile2_decr3 |> mutate(integer1 = as.integer(integer1)))
datafile3_decr3 = decrypt_data(con3, datafile3_encr, vars_to_decrypt); identical(datafile3, datafile3_decr3 |> mutate(integer1 = as.integer(integer1)))


### Combined decryption
datafile123 = datafile1 |> bind_rows(datafile2) |> bind_rows(datafile3)
datafile123_encr = datafile1_encr |> bind_rows(datafile2_encr) |> bind_rows(datafile3_encr)

datafile123_decr1 = decrypt_data(con1, datafile123_encr, vars_to_decrypt); identical(datafile123, datafile123_decr1 |> mutate(integer1 = as.integer(integer1)))
datafile123_decr2 = decrypt_data(con2, datafile123_encr, vars_to_decrypt); identical(datafile123, datafile123_decr2 |> mutate(integer1 = as.integer(integer1)))
datafile123_decr3 = decrypt_data(con3, datafile123_encr, vars_to_decrypt); identical(datafile123, datafile123_decr3 |> mutate(integer1 = as.integer(integer1)))
```



# Number of data points
```{r}
source("C:\\Working projects\\ADM\\C_OpalLUMCADM\\R\\func_number_datapoints.R")

var1 = number_datapoints(datafile, var, cat, count_Missings = FALSE)
var2 = number_datapoints(datafile, var, cat, count_Missings = TRUE)

var1$`Mlstr_area::nr_data_points`
var2$`Mlstr_area::nr_data_points`
```






# End
Below are some things maybe for later


## ADM testdataset2
```{r, eval = FALSE, echo = TRUE}
datafile <- tibble::tribble(
  ~id, ~var1, ~var2, ~var3,          ~var4,        ~var5,                 
  "1", 24L,    8.3,  "test",         "2024-01-01", "2020-01-01 12:00",   
  "2", 0L,     7.9,  "test2",        NA,           NA,                   
  "3", 1L,     NA,   "test?3",       "2995-01-01", "2995-01-01 01:01:01",
  "4", 3L,     79,   NA,             "1923-07-12", "2023-07-12",          
  "5", -99L,   120,  "test$5",       "2021-02-05", "1980-05-02 12:02")


var <- tibble::tribble(
  ~name,  ~valueType, ~`label:en`,  ~unit,    ~index, ~min,         ~max,
  "var1", "integer",  "var1 label", "",       1,      "",           "",
  "var2", "decimal",  "var2 label", "mmol/l", 2,      "0",          "100",
  "var3", "text",     "var3 label", "",       3,      "" ,          "" ,
  "var4", "date",     "var4 label", "",       4,      "1980-01-01", "2023-01-01",
  "var5", "datetime", "var5 label", "",       5,      "",           "2000-01-01 00:00:00")


cat <- tibble::tribble(
  ~variable, ~name,                 ~missing, ~`label:en`,
  "var1",     "1",                  0,        "Option1",
  "var1",     "0",                  0,        "Option0",
  "var1",    "-99",                 1,        "Unknown",
  "var1",    "-98",                 1,        "Not applicable",
  "var2",    "-99",                 1,        "Unknown",
  "var2",    "-98",                 1,        "Not applicable",
  "var4",    "2995-01-01",          1,        "Unknown",
  "var5",    "2995-01-01 01:01:01", 1,        "Unknown",
  "var5",    "2998-01-01 00:00:00", 1,        "Not applicable")
  
# save(datafile, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_datafile.Rdata")
# save(var, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_var.Rdata")
# save(cat, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_cat.Rdata")
```


## ADM testdataset2_adjust
```{r, eval = FALSE, echo = TRUE}
datafile <- tibble::tribble(
  ~id, ~var1, ~var2, ~var3,          ~var4,        ~var5,                 ~var6,
  "1", 24L,    8.3,  "test",         "2024-01-01", "2020-01-01 12:00",    18.0,
  "2", 0L,     7.9,  "test2",        NA,           NA,                    22.5,
  "3", 1L,     NA,   "test?3",       "2995-01-01", "2995-01-01 01:01:01", 23.0,
  "4", 3L,     79,   NA,             "1923-07-12", "2023-07-12",          21.3, 
  "5", -99L,   120,  "test$5",       "2021-02-05", "1980-05-02 12:02",    21.2, 
  "6", -12L,   -98,  "hypothyroïde", "2002-09-13", "2018-10-01 23:59",    28.0)
  
var <- tibble::tribble(
  ~name,  ~valueType, ~`label:en`,  ~unit,    ~index, ~min,         ~max,
  "var1", "integer",  "var1 label", "",       1,      "",           "",
  "var2", "decimal",  "var2 label", "mmol/l", 2,      "0",          "100",
  "var3", "text",     "var3 label", "",       3,      "" ,          "" ,
  "var4", "date",     "var4 label", "",       4,      "1980-01-01", "2023-01-01",
  "var5", "datetime", "var5 label", "",       5,      "",           "2000-01-01 00:00:00",
  "var6", "decimal",  "BMI",        "",       6,      "10",         "100")
  
cat <- tibble::tribble(
  ~variable, ~name,                 ~missing, ~`label:en`,
  "var1",     "1",                  0,        "Option1",
  "var1",     "0",                  0,        "Option0",
  "var1",    "-99",                 1,        "Unknown",
  "var1",    "-98",                 1,        "Not applicable",
  "var2",    "-99",                 1,        "Unknown",
  "var2",    "-98",                 1,        "Not applicable",
  "var4",    "2995-01-01",          1,        "Unknown",
  "var5",    "2995-01-01 01:01:01", 1,        "Unknown",
  "var5",    "2998-01-01 00:00:00", 1,        "Not applicable")
  
# write.csv(datafile, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_datafile_adjust.csv", quote = TRUE, row.names = FALSE)
# write.csv(var, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_var_adjust.csv", quote = TRUE, row.names = FALSE)
# write.csv(cat, file = "I:\\ADM\\_UITVOER\\OPAL\\Development standard scripts opal\\randomdata_removed-parts-importscript_cat_adjust.csv", quote = TRUE, row.names = FALSE)
```











