---
title: "Example OpalLUMCADM"
author: "Lars van der Burg"
date: "2024-02-27"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(diffdf)
library(openxlsx)

library(yaml)

library(rtres)

library(opalr)
library(OpalLUMCADM)
```


In this Rmarkdown document we show all steps that show how to use OpalLUMCADM for working with Opal.


# Initializations
## Dataset
```{r}
load("example/Example_datafile.RData") |> as_tibble()
load("example/Example_var.RData") |> as_tibble()
load("example/Example_cat.RData") |> as_tibble()
  

datafiles = list("first" = datafile, "second" = datafile, "third" = datafile) 
vars = list("first" = var, "second" = var, "third" = var) 
cats = list("first" = cat, "second" = cat, "third" = cat)
```


## Opal
Make sure to have an active opal session

```{r}
opal_url = "https://opal-acc93.clinicalresearch.nl"
opal_token = keyring::key_get("token_acc_clinicalresearch")

opal = opalr::opal.login(url = opal_url, token = opal_token)


projname = tablename = "Example"
if(isFALSE(opalr::opal.project_exists(opal, projname))){
  opalr::opal.project_create(opal = opal, project = projname, database = "mongodb", title = projname, description = "Example project for OpalLUMCADM", tags = "TEST")
}

tablenames = paste0(tablename, 1:3)
ent = "Participant"
``` 

## Paths
```{r}
path_diffdf = paste0(getwd(), "/example/Diffdf/")
if(!dir.exists(path_diffdf)){
  dir.create(path_diffdf)
}
```


# Basic table statements
## Write table
Basis function just to write a table from R to opal.
```{r write}
opal = opalr::opal.login(url = opal_url, token = opal_token)

write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_basis"), 
                   datafile = datafile, var = var, cat = cat, ent = ent, 
                   action = "write", id.name = "id", max_tries = 2)
```


## Read table
Basis function just to read in a table from opal and make it R ready by discarding some of the attributes
```{r import}
opal = opalr::opal.login(url = opal_url, token = opal_token)

imported_table = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_basis"), 
                                     id.name = "id", max_tries = 2)
```


## diffdf report
Function to obtain the differences between the datafile and dictionaries. There are many different ways this function can be used, check the function documentation.
```{r diffdf}
datafile2 = imported_table$datafile
var2 = imported_table$var
cat2 = imported_table$cat


report_diffdf = check_diffdf_opal_generic(datafile = datafile, datafile2 = datafile2, 
                                          var = var, var2 = var2, cat = cat, cat2 = cat2, 
                                          comparison = "mod", comp_key = "id", date_time_format = c("Ymd", "Ymd HMS"), suppress_warnings = TRUE, 
                                          report_path = path_diffdf, report_name = "Report_diffdf", report_title = "This is a trial run",
                                          opt_repl_castor = TRUE, opt_rm_VarDiff_null = FALSE, aggregate_VarDiff = TRUE, 
                                          opt_calc_VarDiff_spaces  = TRUE, opt_calc_VarDiff_diff = TRUE, opt_rm_VarDiff_diff_0 = FALSE) 
```


## Create table
Function to import a table into opal. 

An optional diffdf report will be made if report_path != FALSE. if report_path = NULL, it will not be saved, just outputted by the function
```{r create}
opal = opalr::opal.login(url = opal_url, token = opal_token)

report_import1 = import_create_table_opal(opal = opal, projname = projname, tablename = tablename, ent = ent, 
                                          datafile = datafile, var = var, cat = cat, 
                                          action = "write", id.name = "id", comparison = "mod",
                                          report_path = path_diffdf, report_name = "Report_import")


report_import2 = import_create_table_opal(opal = opal, projname = projname, tablename = tablename, ent = ent, 
                                          datafile = datafile, var = var, cat = cat, 
                                          action = "update", id.name = "id", comparison = "mod", child_lock = FALSE,
                                          report_path = path_diffdf, report_name = "Report_import_update")


report_import3 = import_create_table_opal(opal = opal, projname = projname, tablename = tablename, ent = ent, 
                                          datafile = datafile, var = var, cat = cat, 
                                          action = "overwrite", id.name = "id", comparison = "mod", child_lock = FALSE,
                                          report_path = path_diffdf, report_name = "Report_import_overwrite")
```


## Move/copy table
Function to move a table in opal.
```{r copy}
opal = opalr::opal.login(url = opal_url, token = opal_token)

report_copy = import_copy_table_opal(opal = opal, projname = projname, tablename = tablename, ent = ent,
                                     opal2 = opal, projname2 = projname, tablename2 = paste0(tablename, "_copy"), 
                                     report_path = path_diffdf, report_name = "Report_copy")
```

Possible to move more than table at once, with a wrapper around `import_copy_table_opal`.
```{r multi copy}
opal = opalr::opal.login(url = opal_url, token = opal_token)

report_copy = import_copy_table_opal_many(opal = opal, projnames = projname, tablenames = rep(tablename, 3), ent = ent,
                                          opal2 = opal, projnames2 = projname, tablenames2 = paste0(tablenames, "_copy"),
                                          report_path = path_diffdf, report_name = "Report_copy_many")
```


## Delete table
Function to easily delete a table from opal
```{r}
opal = opalr::opal.login(url = opal_url, token = opal_token)

delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_copy"), child_lock = FALSE)
```

Only one table can be deleted per statement, a for-loop can be used to delete multiple tables
```{r}
for(tabname in tablenames){
  delete_table_opal(opal = opal, projname = projname, tablename = paste0(tabname, "_copy"), child_lock = FALSE)
}
```


## 
### Dictionary update
It is possible to update the dictionary, there are a few things important:
- Columns can be added to the var dictionary, for example the encrypted column
- Labels can be added or altered of *existing* variables only
- The entityType needs to be supplied to the var dictionary (if it is not the default Participant)

```{r}
var_adj = var |> select(-"encrypted")
cat_adj = cat |> slice(-5)  # level 2 of integer2 is gone


write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_dict"), 
                   datafile = datafile, var = var_adj, cat = cat_adj, ent = ent, 
                   action = "write", id.name = "id", max_tries = 2)

imported_table1 = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_dict"), id.name = "id", max_tries = 2)

"encrypted" %in% colnames(imported_table1$var4copy)
imported_table1$cat4copy |> filter(variable == "integer2")


cat_adj = cat_adj |> 
  mutate(`label:en` = ifelse(variable == "integer2" & name == "3", "level 3a", `label:en`)) |> 
  bind_rows(bind_cols(variable = "integer2", name = "2", missing = FALSE, `label:en` = "level2"))
var_adj = var_adj |> 
  mutate(encrypted = var$encrypted,
         entityType = ent)  # Required if entityType is not default `Participant`

opalr::opal.table_dictionary_update(opal, projname, paste0(tablename, "_dict"), var_adj, cat_adj)


imported_table2 = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_dict"), id.name = "id", max_tries = 2)

"encrypted" %in% colnames(imported_table2$var4copy)
imported_table2$cat4copy |> filter(variable == "integer2")





delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_dict"), child_lock = FALSE)
```

## Adjust script
It can happen that you just want to adjust your table or dictionary, and do not want to completely re-do everything. Here is a script exploring some options.

It is not a function. Because the possible adjustments are infinite, it is very difficult to build a comprehensive function for that. But you can simply use the `import_table_opal2R` and `write_table_R2opal` functions. Example is only run with dataset = 2.


Hier bevindingen:
- Rijen kunnen gemakkelijk los toegevoegd worden, een var dictionary is wel essentieel
- Overwrite vs Force. Als dataset al bestaat en wil updaten met nieuwe rij:
  - Overwrite = FALSE & Force = FALSE: error and nothing happens (`Destination table will be updated. There could be data dictionary conflicts. Use 'force' parameter to proceed.`)
  - Overwrite = TRUE & Force = FALSE: error and nothing happens (`Destination table needs to be deleted or truncated. Use 'force' parameter to proceed.`)
  - Overwrite = FALSE & Force = TRUE: updated table: example dataset including the extra row
  - Overwrite = TRUE & Force = TRUE: first delete table and make new table: example dataset only containing the extra row)
  - Conclusie is dus als volgt: overwrite bepaalt wat er gebeurt (FALSE = update; TRUE = overwrite), maar de overwrite actie wordt alleen uitgevoerd als de force parameter = TRUE. Wanneer force = FALSE kan je niet updaten of overwriten.
- Kolommen kunnen ook toegevoegd worden, de var dictionary mag dan wel alleen informatie over de nieuwe kolom bevatten
- Waardes kunnen dan weer wel los aanpepast worden, alleen id en kolom met waarde nodig
- Dictionary kan upgedate worden, entityType in var dictionary is essentieel (anders error (als entityType niet de default `Participant` is))


### Data from opal
```{r}
opal = opalr::opal.login(url = opal_url, token = opal_token)



## To make a separate table for each setting
import_copy_table_opal(opal = opal, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_add_row"))
import_copy_table_opal(opal = opal, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_add_col"))
import_copy_table_opal(opal = opal, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_adjust_val"))
import_copy_table_opal(opal = opal, projname = projname, tablename = tablename, report_path = FALSE,
                       projname2 = projname, tablename2 = paste0(tablename, "_remove_val"))



imported_table = import_table_opal2R(opal = opal, projname, tablename)

datafile4copy = imported_table$datafile4copy
var4copy = imported_table$var4copy
cat4copy = imported_table$cat4copy



## Add a new row 
new_row = tibble::tribble(
  ~id,       ~group, ~integer1, ~text1,      ~date1,        ~datetime1,
  "Sample0", "A",    -40,       "Some text", "2002-09-13", "2018-10-01 23:59:00"
)

write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_add_row"), 
                   datafile = new_row, var = var4copy, cat = cat4copy,
                   ent = ent, action = "update", child_lock = FALSE)
imported_table_add_row = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_add_row"))
imported_table_add_row$datafile4copy |> nrow()


    
## Add a new column
new_col = tibble("integer0" = round(rnorm(nrow(datafile4copy), sd = 5)))
new_col_var = bind_cols(name = "integer0", valueType = "integer", `label:en` = "newly added variable")

var4copy |> bind_rows(new_col_var)

write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_add_col"), 
                   datafile = datafile4copy |> bind_cols(new_col) |> select(id, integer0), var = new_col_var, cat = cat4copy, 
                   ent = ent, action = "update", child_lock = FALSE)
imported_table_add_col = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_add_col"))
imported_table_add_col$datafile4copy
  

  
## Due to checking labjournal a missing value can be replaced
datafile4copy = datafile4copy |> mutate(integer3 = if_else(id == "Sample14", 79, integer3))
datafile4copy = datafile4copy |> mutate(decimal2 = if_else(id == "Sample11", NA, decimal2))
  
write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_adjust_val"),
                   datafile = datafile4copy |> filter(id == "Sample14"), var = var4copy, cat = cat4copy, 
                   ent = ent, action = "update", child_lock = FALSE)
imported_table2_Val1 = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_adjust_val"))
imported_table2_Val1$datafile4copy |> filter(id == "Sample14")

  
write_table_R2opal(opal = opal, projname = projname, tablename = paste0(tablename, "_remove_val"), 
                   datafile = datafile4copy |> filter(id == "Sample11"), var = var4copy, cat = cat4copy, 
                   ent = ent, action = "update", child_lock = FALSE)
imported_table2_Val2 = import_table_opal2R(opal = opal, projname = projname, tablename = paste0(tablename, "_remove_val"))
imported_table2_Val2$datafile4copy |> filter(id == "Sample11")



delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_add_row"), child_lock = FALSE)
delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_add_col"), child_lock = FALSE)
delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_adjust_val"), child_lock = FALSE)
delete_table_opal(opal = opal, projname = projname, tablename = paste0(tablename, "_remove_val"), child_lock = FALSE)
```


# Views
The idea behind opal is that instead of sharing your actual datafile, you can share a view with others. This is an up-to-date copy of the actual datafile. 

With `make_opal_view` you can make a view.
```{r}
opal = opalr::opal.login(url = opal_url, token = opal_token)

view_report = make_opal_view(opal = opal, projname = projname, tablename = tablename, 
                             opal_view = opal, projname_view = projname, tablename_view = paste0(tablename, "_view"),
                             var = var, cat = cat, update = FALSE, comp_key = "id", ent = ent, 
                             report_path = path_diffdf, report_name = "Report_view")
```

Note: if you adjust the underlying table (e.g., remove a variable that the view uses) the view will become deprecated.

This up-to-date copy can also be adjusted in certain ways, if you want to share something that is (slightly) different from the actual datafile.


## Updating the view
The idea is that you can simply change the directory, and than update the view. That is also possible with `make_opal_view`.

Currently, instead of updating, we replace the view. Because with an update it is not yet possible to downsize the view.
```{r}
var_adj = var; var_adj[var_adj$name == "integer2", "label:en"] = "integer with categories; but with a more fancy label"
cat_adj = cat; cat_adj[cat_adj$variable == "integer2" & cat_adj$`label:en` == "level 1", "label:en"] = "level 1; but fancier label"


view_report_update = make_opal_view(opal = opal, projname = projname, tablename = tablename, 
                                    opal_view = opal, projname_view = projname, tablename_view = paste0(tablename, "_view"),
                                    var = var_adj, cat = cat_adj, update = TRUE, comp_key = "id", ent = ent, child_lock = FALSE,
                                    report_path = path_diffdf, report_name = "Report_view_update")
```


## Subset of the data
Can subset the data in various ways:
- select the variables you require in the dictionary
- specify an entityFilter with an evaluable statement. Currently, the only workable entityFilter in `make_opal_view` is to match a specific level of a variable. Many more entityFilters are possible with own written code.

```{r, eval = FALSE, echo = FALSE}
var_adj = var; var_adj = var_adj |> filter(name == "group" | str_detect(name, "integer"))
cat_adj = cat; cat_adj = cat_adj |> filter(variable %in% var_adj$name)


view_report_subset = make_opal_view(opal = opal, projname = projname, tablename = tablename,
                                    opal_view = opal, projname_view = projname, tablename_view = paste0(tablename, "_view_subset"),
                                    var = var_adj, cat = cat_adj, update = FALSE, comp_key = "id", ent = ent, EntityFilter = c("group" = "B"),
                                    report_path = path_diffdf, report_name = "Report_view_subset")
```

```{}
## More difficicult sources
## https://opaldoc.obiba.org/en/latest/magma-user-guide/methods.html
new_rows = tribble(
  ~name, ~valueType, ~`label:en`, ~unit, ~index, ~min, ~max, ~script,
  "var6", "integer", "duplicate of var1", "", 6, "", "", "$('var1')",
  "var7", "integer", "var1 + 10", "", 7, "", "", "$('var1').plus(10)",
  "var8", "decimal", "var1 * var1", "", 8, "", "", "$('var1').multiply($('var1'))",
  # "var9", "integer", "var7 / var1", "", 9, "", "", "$('var7').div($('var1'))",  ## ERROR IN OPAL: No such variable var7 --> Dus waarschijnlijk kan die alleen berekeningen doen op datafile variabele die bestaan? Of er is een volgorde? Can't get the view to R anymore
  "var10", "text", "var3 not", "", 10, "", "", "$('var3').not('test')",
  "var11", "text", "var1 vs var2", "", 11, "", "", "if($('var1').lt($('var2')).value()){true}else{false}",
  "var12", "integer", "var1 vs var2", "", 13, "", "", "if($('var1').lt($('var2')).value()){1}else{0}",
  "var13", "text", "var1 vs var2", "", 14, "", "", "if($('var1').lt($('var2')).value()){$('var2')}else if($('var1').gt($('var2')).value()){$('var1')}",  ## Var1 < var2 --> var2; var1 > var2 --> var1
  "var14", "text", "isNumeric", "", 15, "", "", "$var('var1').isNumeric()",  # without var --> ERROR IN OPAL: Cannot find function isNumeric
  "var15", "decimal", "type", "", 16, "", "", "$('var2').type('decimal')",
  "var16", "decimal", "type + 10", "", 17, "", "", "$('var2').type('decimal').plus(10)",
  "var17", "text", "Date diff", "", 18, "", "", "$('var4').eq($('var5'))"
)


var_adj2 = var |>
  mutate(script = paste0("$('", name, "')")) |> 
  bind_rows(new_rows)

cat_adj2 = cat


view_report_update = make_opal_view(opal = opal, projname = projname, tablename = paste0(tablename, "_EntFilt"), 
                                    opal_view = opal, projname_view = projname, tablename_view = paste0(tablename, "_EntFilt_view2"),
                                    var = var_adj2, cat = cat_adj2, update = TRUE, comp_key = "id", ent = ent, report_path = NULL)
``` 



# Checks
## All checks
There is a single function to run all checks that we currently want to run. It does not change the data, but just gives you an indication whether or not something MIGHT be wrong. Sometimes things are missing for a reason and is a reported issue not a real problem. Always be alert.
```{r}
report_checks = checks_opal_R(datafile = datafile, var = var, cat = cat, key = "id", min_max = TRUE, silent = FALSE)
```


### min/max check
One of the checks that is done in the `checks_opal_R` function is check_categoriesminmax_generic. It is also possible to run this check separately and to run it for multiple datasets at once (that is not possible with the `checks_opal_R` function, that can only handle a single datafile/dictionary)
```{r}
report_minmax = check_categoriesminmax_generic(datafiles = list("check" = datafile), vars = list("check" = var), cats = list("check" = cat), report_path = NULL)
```

It is also possible to run check_categoriesminmax_generic for multiple datasets at once
```{r}
report_minmax_many = check_categoriesminmax_generic(datafiles = datafiles, vars = vars, cats = cats, report_path = NULL)
```


#### Conditional
With the optional conditional argument one can indicate that min/max should only be checked if a specific condition is met. 

For example, when only the integer3 is checked when integer4 == 1
```{r}
datafile |> filter(group == "B") |> select(id, group, integer3, integer4)

report_minmax$check$min_max |> filter(`variable name` == "integer3")


report_minmax_CondF = check_categoriesminmax_generic(datafiles = list("check" = datafile), vars = list("check" = var), cats = list("check" = cat),
                               conditional = list("integer3 ~ group == 'B'"), includeNA = FALSE, report_path = NULL)
report_minmax_CondF$check$min_max |> filter(`variable name` == "integer3")


report_minmax_CondT = check_categoriesminmax_generic(datafiles = list("check" = datafile), vars = list("check" = var), cats = list("check" = cat),
                               conditional = list("integer3 ~ group == 'B'"), includeNA = TRUE, report_path = NULL)
report_minmax_CondT$check$min_max |> filter(`variable name` == "integer3")
```


### Datafile conform var
The `checks_opal_R` function also checks if the specified valueTypes are in concordance with the datafile types. Two functions are of interest here: 
1) datafile_conform_var_check, which just signals whether or not a mismatch is detected (this is run inside `checks_opal_R`) 
2) datafile_conform_var_change, which actually changes the data classes of the datafile for such a mismatch (this is NOT run inside `checks_opal_R`). So if you want to actually change data classes, you have to run `datafile_conform_var_change` separately. 

#### Check
This function will check two things and give a report with the findings
1) whether the valueTypes are named as Opal wants them
2) whether the class of datafile are compatible with the specified valueType. Conversions can be performed with `datafile_conform_var_change`
```{r}
valueType_check = datafile_conform_var_check(datafile, var)
```


#### Change
This function will convert the datafile class to match with the valueType. If datafile class is a character and valueType is a date or datetime, there is no permanant conversion performed, but we will check the conversion that will be done in opal via the function `lubridate::parse_date_time`. With date_time_format you can specify the formats that are observed in the datafile. 
```{r}
datafile_conversion = datafile_conform_var_change(datafile = datafile, var = var, date_time_format = c("Ymd", "Ymd HMS", "Ymd HM"))
datafile = datafile_conversion$datafile


opaltable = opalr::dictionary.apply(tibble = datafile, variables = var, cat = cat)
```



# Encryption & Decryption
```{r}
vars_to_encrypt = vars_to_decrypt = c("integer1", "text2", "date4")

con = tres_connect(base_url = "https://mereden.msbi.nl/Tres", domain = "dw.clinicalresearch.nl", project = "Lumina", username = "ADMs2s", password = keyring::key_get("TRES_opals2s_lumina"))


datafile_encr = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = FALSE, salted_encryption = FALSE)
datafile_decr = decrypt_data(con = con, data = datafile_encr, vars_to_decrypt = vars_to_decrypt)

all.equal(datafile, datafile_decr |> mutate(integer1 = as.integer(integer1)))


datafile_encr = encrypt_data(con = con, data = datafile, vars_to_encrypt = vars_to_encrypt, search_image = TRUE, salted_encryption = FALSE)
datafile_encr |> mutate(across(all_of(vars_to_encrypt), ~vec_extract_search_image(.x), .names = "{.col}_SI")) |> select(vars_to_encrypt, "integer1_SI", "text2_SI", "date4_SI")
```


# Number of data points
```{r}
source("C:\\Working projects\\ADM\\C_OpalLUMCADM\\R\\func_number_datapoints.R")

var1 = number_datapoints(datafile, var, cat, count_Missings = FALSE)
var2 = number_datapoints(datafile, var, cat, count_Missings = TRUE)

var1$`Mlstr_area::nr_data_points`
var2$`Mlstr_area::nr_data_points`
```


# MICA
Things to keep in mind:
- PublishMica needs to have view permissions on the project that contains the table you want to show in Mica
- spaces (" ") are replaced by underscore ("_") in opal for the index, so make sure there are no spaces in the index (title or description is fine)
- If you add annotations manually to the var dictionary. If something is not applicable, don't write NA or NULL, but write it with ""

```{r}
taxonomy = openxlsx::read.xlsx("Mica/Taxonomy_example.xlsx", sheet = "Taxonomy")
annotations = openxlsx::read.xlsx("Mica/Taxonomy_example.xlsx", sheet = "Annotations")

## File required to create a taxonomy in Opal
YAML = taxonomy_setup(taxonomy)
writeLines(YAML, "Mica/Taxonomy_example.yml")

## Update the annotations
write_table_R2opal(opal, projname, paste0(tablename, "_Mica"), datafile = datafile, var = var, cat = cat, ent = ent, action = "write")
opalr::opal.annotate(opal, projname, paste0(tablename, "_Mica"), annotations)


datadict = OpalLUMCADM::import_table_opal2R(opal, projname, paste0(tablename, "_Mica"))
datadict$var4copy |> select(name, starts_with("Tax_example"))
``` 


Same files that need to be supplied can be obtained from Opal/Mica

```{}
library(micar)
mica = micar::mica.login(url = "https://mica-acc93.clinicalresearch.nl/")
taxonomy_mica = micar::mica.taxonomies(mica)


annotations_opal = opalr::opal.annotations(opal, projname, paste0(tablename, "_Mica"))

## YAML can best be downloaded manually from opal
opalr::opal.taxonomies(opal)
``` 


